\chapter{Expressions and Operators}
	\section{Assignment}
		Assignment is done with \texttt{=}. As mentioned above, variables and declared with the \texttt{varname vartype} syntax. Variables can be assigned to a single value or to the result of an expression. 
		
		\begin{lstlisting}
x float = 4.0;
y int = 5/2 + 1; 	/* y = 3 */
		\end{lstlisting}
		
		\noindent Array assignment is done with Java-like syntax. Note that the size of the array must be specified in the declaration. 
		
		\begin{lstlisting}
x int[5] = {0,1,2,3,4};
		\end{lstlisting}
		
		\noindent Pointer types are denoted with a \texttt{*} which must be attached to the primitive type that they reference.
		
		\begin{lstlisting}
x int  = 4;   /* x = 4  */
y int* = &x;  /* *y = 4 */
		\end{lstlisting}
		
	\section{Arithmetic Operations}
		Democritus supports all the arithmetic operations standard to most general-purpose languages like C and Java. Note that casting is not built into the language; this functionality will instead be implemented through the standard library.
		
		\subsection{Addition and Subtraction}
			Addition works with the \texttt{+} character, behaving as expected.
			\begin{lstlisting}
x int = 4;
y int = 2;
x = x+y; 	/* x = 6  */
y = y-x 	/* y = -4 */
			\end{lstlisting}
						
		\subsection{Multiplication}
			Multiplication follows the same rules as well. 
			\begin{lstlisting}
x int = 4;
y int = 2;
x = x*y+y; /* x = 10 */
			\end{lstlisting}
			

		\subsection{Division}
			Democritus will default to integer division, unless both types provided are floats. 
			\begin{lstlisting}
x int = 5;
y int = 2;
x = x/y; 	/* x = 2  */

a float = 4.0;
b float = 2.0;
a = b/a; 	/* a = 2.0 approximately */
			\end{lstlisting}

        \subsection{Modulus}
            The remainder of an integer division operation can be accessed via the modulo \texttt{\%} operator. 
            \begin{lstlisting}
x int = 8;
y int = 5;
x = x%y; 	/* x = 3  */
			\end{lstlisting}
	
		\subsection{Bit Shifting}
			Integers can be bit-shifted with \texttt{\textgreater\textgreater} and \texttt{\textless\textless}.
			\begin{lstlisting}
x int = 9;
y int = x>>1;  /* y = 4  */
x = y<<2; 	   /* x = 16 */
			\end{lstlisting}
		
	\section{Boolean Expressions}
		Democritus features all of the standard logical operators, following Java-style syntax. Each expression will return a boolean value of true or false.

        \subsection{Equality}
            Equality is tested with the \texttt{==} operator. Inequality is tested with \texttt{!=}. 
            \begin{lstlisting}
x int = 8;
y int = 8;
z boolean = (x==y);  	/* z = true  */
z = (x==(y+1));         /* z = false */
z = (x!=(y+1));         /* z = true  */
			\end{lstlisting}
			
		\subsection{Negation}
			Negation is done with \texttt{!}, a unary operation.

        \subsection{Comparison}
            Democritus also features the \texttt{\textless}, \texttt{\textless =}, \texttt{\textgreater}, and \texttt{\textgreater =} operators.  
            
            \begin{lstlisting}
x int = 9;
y int = 8;
z int = 8;

x>y;    /* true */
y>=z;   /* true */
z<x;    /* true */
            \end{lstlisting}

        \subsection{Chained Expressions}
            Boolean expressions can be chained with \texttt{\&\&} and \texttt{\textbar\textbar}, representing \texttt{and} and \texttt{or}, respectively. These operators have lower precedence than any of the other boolean operators described above. The \texttt{and} operator has a higher precedence than \texttt{or}.

            \begin{lstlisting}
x int = 9;
y int = 8;
z int = 8;

(x>y && y<x);    		/* false */
(x>y || y<x);  	 		/* true */
(x>y && y<x || z==y)    /* true */
            \end{lstlisting}
            
	\section{Pointers and References}
		Pointers and dereferencing utilizes the same syntax as C. The unary operator \texttt{\&} gives a variables address in memory, and the operator \texttt{*} dereferences a pointer. See the assignment section for usage.
		
	\section{Array access}
		Array access is done with \texttt{[\textit{i}]} where \textit{i} is the index being accessed. 
		
		\begin{lstlisting}
x int[5] = {0,1,2,3,4};
y int = x[2];  			/* y = 2 */
		\end{lstlisting}
		